Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Correcting timeseries to time series
Following advice online:

https://english.stackexchange.com/a/570738

---
title: Think deep for reef resilience
author:
  - name: Benjamin Grier
    email: grierben777@gmail.com
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          state: QLD
    attributes:
        corresponding: true
  - name: Takuya Iwanaga
    email: t.iwanaga@aims.gov.au
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          city: Cape Cleveland
          state: QLD
    attributes:
        corresponding: true
abstract: |
    Coral reefs globally face unprecedented threats from climate change, with the Great Barrier
    Reef (GBR) experiencing severe declines in coral cover from thermal stress events.
    Understanding drivers behind reef resilience to climate impacts is critical for conservation
    planning and intervention strategies. A scenario-based population modelling approach was
    adopted with larval connectivity dynamics, thermal adaptation, and environmental factors
    to assess coral reef resilience across GBR locations under future climate projections.
    Using five Global Climate Models (GCMs) spanning five Equilibrium Climate Sensitivities,
    we analysed reef cover timeseries and the ability of reefs to maintain positive carbonate
    production budgets under future conditions. Contrary to expectations, larval connectivity
    patterns did not correlate with increases in maintenance of positive carbonate budgets.
    Instead, reef depth emerged as the primary predictor of cover-timeseries resilience and
    maintenance of positive carbonate budgets, with deeper reefs benefiting from thermal
    refugia that reduce bleaching impacts. Our findings suggest that depth provides critical
    protection against thermal stress, contrasting current restoration practices that primarily
    target easily accessible shallow sites. These results have important implications for
    reef management, indicating that depth refugia should be prioritized for further
    consideration in conservation planning alongside connectivity considerations to maximize
    coral survival under continuing climate change.
keywords:
  - Coral reef
  - Climate change
  - Ecosystem resilience
  - Reef depth
  - Reef connectivity
date: last-modified
format:
  elsevier-docx:
    journal:
      name: Journal Name
      formatting: preprint
      cite-style: authoryear
bibliography: references.bib
engine: julia
crossref:
  custom:
  - kind: float
    reference-prefix: SFigure
    key: sfig
editor:
  markdown:
    wrap: 92
---

@@

![Degree Heating Week timeseries (in °C - Weeks) for the GCMs selected. Colours represent
the different management areas across the Great Barrier Reef. Green represents
Cairns-Cooktown, orange represents Far-Northern, blue represents Mackay-Capricorn and red
represents the Townsville-Whitsunday management area. Solid lines represent the
management-area median timeseries and bands represent the 95th percentile confidence
intervals.](../figs/methods_dhw_timeseries.png){#fig-methods-dhws fig-align="center"
width="14.82cm"}

@@

ADRIA-CoralBlox was run for 4096 scenarios for each of the five GCMs (20,480 total model
runs). Here, a scenario represents a unique combination of assumed environmental (namely one
of the five DHW trajectories) and ecological factors including coral growth and mortality
rates, coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol'
sampling method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi
Monte Carlo approach which balances scenario exploration with improved convergence
properties compared to pseudo-random sampling. A detailed comparative analysis of the
sampling methods can be found in [@kucherenko2015]. For the purpose of scenario assessment,
the resulting timeseries for each model run were scaled by the total area of the reef to
represent cover as a percentage of the total estimated reef area. These timeseries of coral
cover were then clustered based on their temporal behaviour.

### Timeseries clustering

@@

Partial Dependence Analysis was conducted to understand the relationship between individual
features and time series cluster predictions using a trained Random Forest (RF) classifier
[@friedman2001]. The analysis utilised the reef characteristics of interest as the
predictors, including the median reef depth, mean DHW for the given GCM, total connectivity
strength, and the outgoing-to-incoming connectivity ratio (\$O:I\$). The full dataset
consisted of the timeseries cluster assignment of each of the 2429 reefs under the five GCMs
explored, for a total of 12,145 samples. The dataset was randomly split into training (60%)
and testing (40%) subsets to ensure model generalisability. The RF classifier achieved a 99%
accuracy on the training set and 66% on the test set.

@@

Using median coral cover timeseries for each reef we calculated the number of years each
reef maintains coral cover above a given percent threshold. The threshold represents the
level of live coral cover necessary to maintain a positive carbonate budget (denoted here as
$\theta_{0}$ to indicate the budget equilibrium). Greater number of years above $\theta_{0}$
therefore suggests greater resilience and opportunity for recovery.

@@

Under each GCM, in all management areas, low, medium and high cluster reefs have low levels
of distinction between DHW trajectories (@sfig-gcm-dhw-trajectories). The highest peak DHW
levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs
timeseries occur under ACCESS-CM2 [@fig-methods-dhws].

![Reef area-relative coral cover timeseries clustered within each management area, for each
GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2060 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2060 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 95th percentile confidence intervals. Each management area is represented five
times (for each GCM), with each GCM displayed
horizontally.](../figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
fig-align="center" width="14.82cm"}

@@

Across many bioregions timeseries clustering identified a ‘low’ cluster containing reefs
that suffer larger declines and generally have lower levels of coral cover than the other
clusters, a ‘medium’ cluster with intermediate levels of coral cover, and a ‘high’ cluster
with greater levels of cover recovery, with persistence to 2050 in some bioregions
(@sfig-access-esm1-5-bioregion-timeseries).

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW timeseries
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean timeseries DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

@@

Our findings suggest that reefs in ‘high’ cover clusters do not have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024; @lachs2023], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The low
importance of thermal tolerance on cover outcomes may be due to reefs experiencing high
levels of thermal stress early in projected timeseries. High - early timeseries thermal
stress events may prevent the intergenerational adaptation of corals to thermal stress and
prevent thermal tolerance from reaching levels greater than 9°C-weeks. **\[TI note: This
paragraph needs to be revisited\]**

@@

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2060 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2060 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../figs/GCM_dhw_timeseries_plot.png){#sfig-gcm-dhw-trajectories
fig-align="center" width="14.82cm"}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2075. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef base-10 log of
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. The colour of each point represents the base-10 log of scaled reef total
connectivity strength. Displays results from median timeseries using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate
fig-align="center" width="14.82cm"}

@@

![Reef area-relative coral cover timeseries clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2060 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2060 within each
bioregion. Solid lines represent the cluster median timeseries and bands represent the 95th
percentile confidence intervals. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/bioregion_cover_timeseries.png){#sfig-access-esm1-5-bioregion-timeseries
fig-align="center" width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster and bands
represent the distribution of correlation levels within each cluster. Displays results for
ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwcor
fig-align="center" width="14.82cm"}

![Mean timeseries DHW values for reefs within each bioregion, separated into clusters
assigned by timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2060
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2060 within each bioregion. Solid lines represent the
median DHW level for each cluster and bands represent the distribution of mean DHW levels
within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwmean
fig-align="center" width="14.82cm"}

![Mean timeseries DHW tolerance value, from 2035 to 2075, for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines represent the median DHW tolerance value for each cluster and bands represent the
distribution of DHW-tolerance levels within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwtol
fig-align="center" width="14.82cm"}

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green)
are clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines represent the median proportion of habitable area occupied for each cluster and bands
represent the distribution of occupied area values within each cluster. Displays results for
ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-esm1-5-bioregion-initialproportion
fig-align="center" width="14.82cm"}

![Source-to-sink connectivity ratios (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median source-to-sink ratio for each cluster and
bands represent the distribution of source-to-sink ratios within each cluster. The vertical
grey line represent a source-to-sink ratio of 0 (in base-10 log scale) indicating equal
outgoing and incoming connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
fig-align="center" width="14.82cm"}

![Total connectivity strength (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by timeseries clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median base-10 log of total connectivity strength for
each cluster and bands represent the distribution of connectivity strength values within
each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity
fig-align="center" width="14.82cm"}

>>>>

---
title: Think deep for reef resilience
author:
  - name: Benjamin Grier
    email: grierben777@gmail.com
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          state: QLD
    attributes:
        corresponding: true
  - name: Takuya Iwanaga
    email: t.iwanaga@aims.gov.au
    affiliations:
        - id: AIMS
          name: Australian Institute of Marine Science
          address: Cape Cleveland Rd
          city: Cape Cleveland
          state: QLD
    attributes:
        corresponding: true
abstract: |
    Coral reefs globally face unprecedented threats from climate change, with the Great Barrier
    Reef (GBR) experiencing severe declines in coral cover from thermal stress events.
    Understanding drivers behind reef resilience to climate impacts is critical for conservation
    planning and intervention strategies. A scenario-based population modelling approach was
    adopted with larval connectivity dynamics, thermal adaptation, and environmental factors
    to assess coral reef resilience across GBR locations under future climate projections.
    Using five Global Climate Models (GCMs) spanning five Equilibrium Climate Sensitivities,
    we analysed reef cover time series and the ability of reefs to maintain positive carbonate
    production budgets under future conditions. Contrary to expectations, larval connectivity
    patterns did not correlate with increases in maintenance of positive carbonate budgets.
    Instead, reef depth emerged as the primary predictor of cover-time series resilience and
    maintenance of positive carbonate budgets, with deeper reefs benefiting from thermal
    refugia that reduce bleaching impacts. Our findings suggest that depth provides critical
    protection against thermal stress, contrasting current restoration practices that primarily
    target easily accessible shallow sites. These results have important implications for
    reef management, indicating that depth refugia should be prioritized for further
    consideration in conservation planning alongside connectivity considerations to maximize
    coral survival under continuing climate change.
keywords:
  - Coral reef
  - Climate change
  - Ecosystem resilience
  - Reef depth
  - Reef connectivity
date: last-modified
format:
  elsevier-docx:
    journal:
      name: Journal Name
      formatting: preprint
      cite-style: authoryear
bibliography: references.bib
engine: julia
crossref:
  custom:
  - kind: float
    reference-prefix: SFigure
    key: sfig
editor:
  markdown:
    wrap: 92
---

@@

![Degree Heating Week time series (in °C - Weeks) for the GCMs selected. Colours represent
the different management areas across the Great Barrier Reef. Green represents
Cairns-Cooktown, orange represents Far-Northern, blue represents Mackay-Capricorn and red
represents the Townsville-Whitsunday management area. Solid lines represent the
management-area median time series and bands represent the 95th percentile confidence
intervals.](../figs/methods_dhw_timeseries.png){#fig-methods-dhws fig-align="center"
width="14.82cm"}

@@

ADRIA-CoralBlox was run for 4096 scenarios for each of the five GCMs (20,480 total model
runs). Here, a scenario represents a unique combination of assumed environmental (namely one
of the five DHW trajectories) and ecological factors including coral growth and mortality
rates, coral thermal tolerances. Scenarios were generated using the low-discrepancy Sobol'
sampling method [@sobol1967; @sobol1993] with Owen Scrambling [@owen1995], a random-Quasi
Monte Carlo approach which balances scenario exploration with improved convergence
properties compared to pseudo-random sampling. A detailed comparative analysis of the
sampling methods can be found in [@kucherenko2015]. For the purpose of scenario assessment,
the resulting time series for each model run were scaled by the total area of the reef to
represent cover as a percentage of the total estimated reef area. These time series of coral
cover were then clustered based on their temporal behaviour.

### Time series clustering

@@

Partial Dependence Analysis was conducted to understand the relationship between individual
features and time series cluster predictions using a trained Random Forest (RF) classifier
[@friedman2001]. The analysis utilised the reef characteristics of interest as the
predictors, including the median reef depth, mean DHW for the given GCM, total connectivity
strength, and the outgoing-to-incoming connectivity ratio. The full dataset consisted of the
time series cluster assignment of each of the 2429 reefs under the five GCMs explored, for a
total of 12,145 samples. The dataset was randomly split into training (60%) and testing
(40%) subsets to ensure model generalisability. The RF classifier achieved a 99% accuracy on
the training set and 66% on the test set.

@@

Using median coral cover time series for each reef we calculated the number of years each
reef maintains coral cover above a given percent threshold. The threshold represents the
level of live coral cover necessary to maintain a positive carbonate budget (denoted here as
$\theta_{0}$ to indicate the budget equilibrium). Greater number of years above $\theta_{0}$
therefore suggests greater resilience and opportunity for recovery.

@@

Under each GCM, in all management areas, low, medium and high cluster reefs have low levels
of distinction between DHW trajectories (@sfig-gcm-dhw-trajectories). The highest peak DHW
levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs
time series occur under ACCESS-CM2 [@fig-methods-dhws].

![Reef area-relative coral cover time series clustered within each management area, for each
GCM analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2060 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2060 within each bioregion. Solid lines represent the cluster median time series and bands
represent the 95th percentile confidence intervals. Each management area is represented five
times (for each GCM), with each GCM displayed
horizontally.](../figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
fig-align="center" width="14.82cm"}

@@

Across many bioregions time series clustering identified a ‘low’ cluster containing reefs
that suffer larger declines and generally have lower levels of coral cover than the other
clusters, a ‘medium’ cluster with intermediate levels of coral cover, and a ‘high’ cluster
with greater levels of cover recovery, with persistence to 2050 in some bioregions
(@sfig-access-esm1-5-bioregion-timeseries).

@@

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW time series
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean time series DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

@@

Our findings suggest that reefs in ‘high’ cover clusters do not have greater thermal
tolerance than reefs in lower coral cover clusters. This is in contrast with literature
suggesting corals have the capacity to adapt to increasing climate change pressures
[@humanes2022; @lachs2024; @lachs2023], including the investigation by Bairos-Novak et al.
[-@bairos-novak2021] that revealed high heritability of bleaching and survival response
traits, used to inform thermal tolerance trait distributions in this investigation. The low
importance of thermal tolerance on cover outcomes may be due to reefs experiencing high
levels of thermal stress early in projected time series. High - early time series thermal
stress events may prevent the intergenerational adaptation of corals to thermal stress and
prevent thermal tolerance from reaching levels greater than 9°C-weeks. **\[TI note: This
paragraph needs to be revisited\]**

@@

![Reef Degree Heating Week time series clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2060 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2060 within each bioregion. Solid lines represent the cluster median time series and bands
represent the 2.5 to 97.5 percentile confidence intervals. Each management area is
represented five times (for each GCM), with each GCM displayed
horizontally.](../figs/GCM_dhw_timeseries_plot.png){#sfig-gcm-dhw-trajectories
fig-align="center" width="14.82cm"}

@@

![Number of years reef coral cover exceeds the carbonate budget threshold level from 2025 to
2075. The carbonate budget thresholds assessed range from 10 to 20% of total reef area.
Bracketed x-axis values represent the Spearman's correlation between reef base-10 log of
total connectivity strength and the number of years cover exceeds the carbonate budget
threshold. The colour of each point represents the base-10 log of scaled reef total
connectivity strength. Displays results from median time series using
ACCESS-ESM1-5.](../figs/ACCESS-ESM1-5/log_total_strength_carbonate_scatter.png){#sfig-access-esm1-5-conn-carbonate
fig-align="center" width="14.82cm"}

@@

![Reef area-relative coral cover time series clustered within each bioregion. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2060 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2060 within each
bioregion. Solid lines represent the cluster median time series and bands represent the 95th
percentile confidence intervals. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/bioregion_cover_timeseries.png){#sfig-access-esm1-5-bioregion-timeseries
fig-align="center" width="14.82cm"}

![Total coral cover – DHW Pearson’s correlation levels for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines represent the median cover – DHW correlation level for each cluster and bands
represent the distribution of correlation levels within each cluster. Displays results for
ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/dhw_cover_cor_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwcor
fig-align="center" width="14.82cm"}

![Mean time series DHW values for reefs within each bioregion, separated into clusters
assigned by time series clustering. ‘Low’ clusters (green) are clusters with the lowest
median area-relative cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters
(orange) are clusters with intermediate median area-relative cover between 2030 and 2060
within each bioregion. ‘High’ clusters (blue) are clusters with the highest median
area-relative cover between 2030 and 2060 within each bioregion. Solid lines represent the
median DHW level for each cluster and bands represent the distribution of mean DHW levels
within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/mean_dhw_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwmean
fig-align="center" width="14.82cm"}

![Mean time series DHW tolerance value, from 2035 to 2075, for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines represent the median DHW tolerance value for each cluster and bands represent the
distribution of DHW-tolerance levels within each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/dhw_tolerance_bioregion_violin.png){#sfig-access-esm1-5-bioregion-dhwtol
fig-align="center" width="14.82cm"}

![Initial proportion of coral habitable area occupied by corals for reefs within each
bioregion, separated into clusters assigned by time series clustering. ‘Low’ clusters
(green) are clusters with the lowest median area-relative cover between 2030 and 2060 within
each bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median
area-relative cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are
clusters with the highest median area-relative cover between 2030 and 2060 within each
bioregion. Solid lines represent the median proportion of habitable area occupied for each
cluster and bands represent the distribution of occupied area values within each cluster.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/initial_proportion_bioregion_violin.png){#sfig-access-esm1-5-bioregion-initialproportion
fig-align="center" width="14.82cm"}

![Source-to-sink connectivity ratios (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median source-to-sink ratio for each cluster and
bands represent the distribution of source-to-sink ratios within each cluster. The vertical
grey line represent a source-to-sink ratio of 0 (in base-10 log scale) indicating equal
outgoing and incoming connectivity levels. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/so_to_si_bioregion_violin.png){#sfig-access-esm1-5-bioregion-logsosi
fig-align="center" width="14.82cm"}

![Total connectivity strength (in base-10 log scale) for reefs within each bioregion,
separated into clusters assigned by time series clustering. ‘Low’ clusters (green) are
clusters with the lowest median area-relative cover between 2030 and 2060 within each
bioregion. ‘Medium’ clusters (orange) are clusters with intermediate median area-relative
cover between 2030 and 2060 within each bioregion. ‘High’ clusters (blue) are clusters with
the highest median area-relative cover between 2030 and 2060 within each bioregion. Solid
lines within each violin represent the median base-10 log of total connectivity strength for
each cluster and bands represent the distribution of connectivity strength values within
each cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/log_total_strength_bioregion_violin.png){#sfig-access-esm1-5-bioregion-totalconnectivity
fig-align="center" width="14.82cm"}
